---
title: "Week 1 Discussion"
author: Madi Calbert
date: January 6, 2024
format: html
---

### Load Libraries

```{r}
library(tidyverse)
library(janitor)
library(usdata)
```

### Import fracking data 

```{r}
fracking <- read_csv(here("week1/data/registryupload_1.csv"))
```

### Data exploration 

```{r}
dim(fracking) #tells you how big the df is

str(fracking) #tells you the structure of the df 

view(fracking) #simply look at the df
```


### Clean Fracking Data 

```{r}
fracking_clean <- fracking %>% 
  
  clean_names() %>% #convert column names to snake case
  
  mutate(job_start_date = str_remove(job_start_date, " AM")) %>% #remove "AM" from start date
  
  mutate(datetime_start = mdy_hms(job_start_date)) %>% #new column with datetime objects
  
  mutate(year = year(datetime_start)) %>% #add a year column with year from datetime_start
  
  select(datetime_start, 
         year, 
         state_name, 
         well_name, 
         total_base_water_volume) %>%  #keep only necessary columns
  
  mutate(state_name = str_to_title(state_name)) %>%  #convert obsvs in the state_name column to Title Case
  
  filter(!str_detect(state_name, pattern = "\\?")) %>% #remove any rows where the state name ends in ?
  
  mutate(state_name = ifelse(       #convert any state abbreviations to full names in the state_name column
    test = str_length(state_name) == 2, #b/c the abbreviations are only 2 characters long (ex: Pa)
    yes = abbr2state(state_name),
    no = state_name
  )) %>% 
  
  mutate(state_name = case_when(  # correct any misspelled state names in the state_name
    state_name == "Colordao" ~ "Colorado",
    state_name == "Loiusiana" ~ "Louisiana",
    state_name == "Louisianna" ~ "Louisiana",
    state_name == "Lousiana" ~ "Louisiana",
    state_name == "New Mexcio" ~ "New Mexico",
    state_name == "Norh Dakota" ~ "North Dakota",
    state_name == "Norht Dakota" ~ "North Dakota",
    state_name == "North  Dakota" ~ "North Dakota",
    state_name == "North Dakata" ~ "North Dakota",
    state_name == "North Dakotta" ~ "North Dakota",
    state_name == "Noth Dakota" ~ "North Dakota",
    state_name == "Pennslvania" ~ "Pennsylvania",
    state_name == "Pennsylavania" ~ "Pennsylvania",
    state_name == "Pennsylvanya" ~ "Pennsylvania",
    state_name == "Penssylvania" ~ "Pennsylvania",
    state_name == "Texasa" ~ "Texas",
    state_name == "Texs" ~ "Texas", 
    state_name == "West Viginia" ~ "West Virginia",
    state_name == "Wyominng" ~ "Wyoming", 
    TRUE ~ state_name # copy over rest of state names from as-is
  )) %>% 
  
  filter(!state_name %in% # remove any rows that do not have a true US state name
           c("Beaver", "Beckham", "Harper", "Hemphill", "Midland", 
             "Red River", "Roosevelt", "Rusk", "State", "Ward"))
  


fracking_clean
```

### NEED TO FINISH 

k. add a column named state_abb with just US state abbreviations
l. move the state_abb column so that it sits immediately after the state_name column
m. rename total_base_water_volume as total_base_water_volume_gal for clarity
n. remove any observations that don’t have a measurement for total_base_water_volume_gal (i.e. NA)
o. reorder rows from earliest to most recent datetime_start values