---
title: "week2-discussion"
author: Madi Calbert
date: January 14, 2024
format: html
---

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                                    setup                                 ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

#..........................load packages.........................
library(tidyverse)

#..........................import data...........................
drought <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2021/2021-07-20/drought.csv')

##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##                            wrangle drought data                          ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

drought_clean <- drought |>

  # select cols of interest & update names for clarity (as needed) ----
  select(date = valid_start, state_abb, drought_lvl, area_pct) |> 

  # add year, month & day cols using {lubridate} fxns ----
  # NOTE: this step isn't necessary for our plot, but I'm including as examples of how to extract different date elements from a object of class `Date` using {lubridate} ----
  mutate(year = year(date),
         month = month(date, label = TRUE, abbr = TRUE),
         day = day(date)) |>

  # add drought level conditions names ----
  mutate(drought_lvl_long = factor(drought_lvl,
                            levels = c("D4", "D3", "D2", "D1","D0", "None"),
                            labels = c("(D4) Exceptional", "(D3) Extreme",
                                       "(D2) Severe", "(D1) Moderate", 
                                       "(D0) Abnormally Dry", 
                                       "No Drought"))) |>
  
  # reorder cols ----
  relocate(date, year, month, day, state_abb, drought_lvl, drought_lvl_long, area_pct)
```
a). Create a stacked area chart (take a look back at lecture 2.3 slides for examples)

b). Figure out how to order the drought level groups in the same way as the USDM version (i.e. level D4 on the bottom, closet to the x-axis, and level D0 at the top) – Hint: explore the position argument

c) Update the colors so that they match the USDM version (use Colorpick Eyedropper to grab HEX codes from the original visualization) – Hint: use scale_fill_manual() to set your new colors

d) Adjust your x-axis “breaks” (i.e. the tick mark values that represent years) – Hint: check out scale_x_date() and scales::breaks_pretty()

e) Adjust your y-axis “breaks” (i.e. the tick mark value that represent percentage of affected area) – Hint: check out scale_y_continous() and scales::label_percent()

f) Remove the “padding” (i.e. space) between your area chart and the x- and y-axes – Hint: check out the expand argument in scale_x_date() and scale_y_continuous()

g) update the plot title

```{r}
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
##       create stacked area plot of CA drought conditions through time     ----
##~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

drought_clean |> 
  
  # b. remove drought_lvl "None" & filter for just CA ----
  filter(drought_lvl != "None",
         state_abb == "CA") %>% 
  ggplot(aes(x = date, y = area_pct, fill = drought_lvl_long)) +
  
  #a. reverse order of groups so level D4 is clsoest to the x axis
  geom_area(position = position_stack(reverse = TRUE)) +
  
  #c. update colors 
  scale_fill_manual(values = c("#853904", "#FF0000", "#FFC100", "#FFD965", "#FFFF00")) +
  
  #d) Adjust your x-axis “breaks” and f) Remove the “padding” 
  scale_x_date(breaks = scales::breaks_pretty(n = 10),
               expand = c(0,0)) +
  
  #e) Adjust your y-axis “breaks” and f) Remove the “padding” 
  scale_y_continuous(breaks = scales::breaks_pretty(n = 10), #increments of 10
                     labels = scales::label_percent(scale = 1), #adds the percent sign
                     expand = c(0,0)) +
  
  #g) update the plot title
 labs(title = "Drought Area in California", 
      x = NULL,
      y = NULL,
      fill = NULL) +
  
  theme_minimal() +
  
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5))
  
```


